### 2020-02-28 manuscript/supplement.tex
finished rejiiging

\documentclass[rmp, reprint, superscriptaddress, floatfix,amsmath]{revtex4-1}
\usepackage{longtable}
\usepackage{color}
\usepackage{graphicx}  % needed for figures
\usepackage[utf8x]{inputenc}
\usepackage{hyperref}
\usepackage{longtable}
\usepackage{xcolor}
\usepackage{soul}
\usepackage{multirow}
\usepackage{longtable}
\usepackage{xr-hyper}
\newcommand{\forcing}{\varepsilon}

\makeatletter
\newcommand*{\addFileDependency}[1]{% argument=file name and extension
  \typeout{(#1)}
  \@addtofilelist{#1}
  \IfFileExists{#1}{}{\typeout{No file #1.}}
}
\makeatother

\newcommand*{\myexternaldocument}[1]{%
    \externaldocument{#1}%
    \addFileDependency{#1.tex}%
    \addFileDependency{#1.aux}%
}
\myexternaldocument{main}

\definecolor{drab}{rgb}{0.59, 0.44, 0.09}
\newcommand{\Richard}[1]{{\color{drab}Richard: #1}}
\definecolor{celestialblue}{rgb}{0.29, 0.59, 0.82}
\newcommand{\Robert}[1]{{\color{celestialblue}Robert: #1}}
\definecolor{purple}{rgb}{0.459,0.109,0.538}
\definecolor{deepsaffron}{rgb}{1.0, 0.6, 0.2}
\newcommand{\Jan}[1]{{\color{deepsaffron}Jan: #1}}
\newcommand{\Emma}[1]{{\color{purple}Emma: #1}}
\definecolor{green}{rgb}{0.15, 0.6, 0.15}
\newcommand{\Valentin}[1]{{\color{green}Valentin: #1}}

\begin{document}

\title{Supplement: Impact of seasonal forcing on a potential SARS-CoV-2 pandemic}

\setcounter{figure}{0}
\renewcommand{\figurename}{Figure S}
\setcounter{table}{0}
\renewcommand{\tablename}{Table S}
\maketitle
The SIR model introduced in equations \ref{eq:SIR} describes the fraction of susceptible $S(t)$, exposed $E(t)$, and infected individuals $I(t)$ in a population. 
The fraction of recovered individuals $R$ is $R=1 - S - E - I$.

An illustration of the model is presented in figure \ref{fig:schemeSIR} with constant transmissibility ($\forcing=0$ in equation \ref{eq:transmissibility}).

When simplifying the model even further and combining the categories of exposed and infectious into one class of infected individuals, the model can be solved analytically.
The simplified system has a steady state
\begin{equation}
\begin{split}
S_0 &= \frac{\nu + b}{\beta} = \frac{1}{R_0} \\
I_0 & = \frac{b(R_0-1)}{\beta}
\end{split}
\label{eq:fixed_point}
\end{equation}
where $\nu$ now is the inverse of average time from exposure to recovery.
This steady state exists when the parameter $R_0 > 1$. For cases where $R_0 < 1$ the epidemic dies out as each infected person generate less than one subsequent infection.

\begin{figure}[htb]
	\centering
	\includegraphics[width=0.9\columnwidth]{figures/SEIR_sketch.pdf}
	\caption{Illustration of the SIER model used with its parameters.}\label{fig:schemeSIR}
\end{figure}

For our analysis of seasonal CoVs, it is instructive to study the stability of this steady state and show that the quantities $[S,I]$ tend to show damped oscillations around the fixed points in Eq.~\ref{eq:fixed_point} with a period
\begin{equation}
T = \frac{4\pi}{\sqrt{4b(\nu(R_0-1) - b) - R_0^2b^2}}
\label{eq:period}
\end{equation}
This behavior is intrinsic to the system and holds whenever transmissibility is constant.

SIR models can show resonance phenomena when the period is about one or two years \citep{dushoff_dynamical_2004,chen_regular_2017}.
Resonance is less likely when the period is much longer than the the period of annual forcing.
Fig.~\ref{fig:intrinsicPeriod} shows the intrinsic period for the parameter range considered and two durations of immunity (left: 10 years, right: 5 years). 
Our parameters choices for $\nu$ $\beta$ lie in the center of the graphs in Fig.~\ref{fig:intrinsicPeriod}. 
For $b=0.1/y$, we expect an intrinsic period is around 3 years, suggesting that resonance is unlikely and if it occurs would lock into 3 year patters.
For $b=0.2/y$, the intrinsic period is closer to 2 years and biennial resonance is possible. 
The data, however, don't suggest strong biennial patterns such that we don't think resonance is a major contributor to seasonality.

In a fully susceptible population, the number of infected individuals initially grows as $I(t)\sim e^{(\beta-\nu-b)t} = e^{(R_0-1)t/(\nu+b)}$.
The doubling time is $\tau_2 = (\nu+b)\log(2)/(R_0-1)$.
Since $b\ll \nu$, we can safely neglect $b$ in the early phase of the pandemic, but it is important when interpreting endemic seasonal CoV data.

\begin{figure*}[htb]
	\centering
	\includegraphics[width=0.9\columnwidth]{figures/Period_phase_space_b_0.1.pdf}
	\includegraphics[width=0.9\columnwidth]{figures/Period_phase_space_b_0.2.pdf}
\caption{Length of the intrinsic period of the SIR model for $\beta$ and $\nu$ values in the range explored in this paper. The duration of immunity is assumed to be $10$ years in the panel on the left and 5 years in the panel on the right. The dark blue region corresponds to $R_0<1$. }
	\label{fig:intrinsicPeriod}
\end{figure*}

\subsection*{Migration between subpopulations}

In the model for seasonal CoVs, we account for the exchange of CoVs with the rest of the world through an additional transition of susceptibles to the exposed category with a rate that is the product of a migation rate and the probability of being exposed abroad. 
Beyond that, no additions to Eq.~\ref{eq:SIR} are necessary.

For the multipopulation models of SARS-CoV-2 pandemics, we model migration as loss terms to each category with a population specific migration rate, followed by redistributions of the pooled migrants to random population according the population specific migration rate. 
Such migration is operating in Figures \ref{fig:nCov_predictions}, \ref{fig:global_3_panel} and \ref{fig:endemic}.

\subsection*{Effect of infection control measures}
The SIR model as formulated in Eq.~\ref{eq:SIR} accounts for seasonality but otherwise assumes the transmission dynamics is constant. 
In reality, individuals change their behavior when they are aware of a contagious disease.
This in particularly true for the current SARS-CoV-2 outbreak where tens of millions of people are under quarantine measures and travel restrictions.
In order to account for such behavioral changes, we add a factor $H$ to equation \ref{eq:SIR} that modulates transmission:
\begin{equation}
\frac{d}{dt} I =  -(\nu+b) I + (1-H)\beta S I + i
\end{equation}
where $H = c\cdot\frac{I^3}{(K^3+I^3)}$ with $c$ the value of the containment parameter. 
The term $H$ is a Hill function of order 3 and inflection point $K$. 
We use this to model a decrease in the rate of new infection smoothly decrease by a factor $c$. 
We used $c=0.5$ and $K=0.03$, implying that once a disease is at a prevalence of 3\%, containment measures reduce transmission by 50\%.
The effect of prevention and control is difficult to assess at present and will certainly vary in time and space.

\begin{figure}
    \centering
    \includegraphics[width=0.5\textwidth]{figures/fraction_positive_HKU1_OC43.pdf}
    \caption{Seasonal variation of HKU1/OC43 CoV positive tests in Stockholm, Sweden. }
    \label{fig:seasonal_variation_supp}
\end{figure}

\begin{figure*}[h]
	\centering
	\includegraphics[width=0.9\textwidth]{figures/scenario_subplot_0.5.pdf}
	\caption{{\bf Model predictions for SARS-CoV-2 case numbers in temperate zones for a pandemic scenario, with varying $\langle R_0\rangle$ and migration.} This is the same plot as shown on the left of Figure \ref{fig:nCov_predictions} with $\forcing=0.5$, however here $\langle R_0\rangle$ varies from 1.3 - 3 and Migration varies from 0.1\% to 10\% per year. Higher migration rates (towards the bottom of the figure) result in earlier introductions and higher likelihood of a peak in early 2020.
	The scenarios with low $\langle R_0\rangle$ are clearly counter-factual. 
	Similarly, higher $\langle R_0\rangle$ (towards the right of the figure) results in more rapid growth and a higher likelihood of an early peak.}
	\label{fig:scenarios_supp}
\end{figure*}

\begin{figure*}[h]
	\centering
	\includegraphics[width=0.9\textwidth]{figures/scenario_subplot_0.15.pdf}
	\caption{{\bf Model predictions for SARS-CoV-2 case numbers in temperate zones for a pandemic scenario, with varying $R_0$ and migration.} This is the same plot as shown on the left of Figure \ref{fig:scenarios_supp} but with $\forcing=0.15$. All scenarios result in a single peak.}
	\label{fig:scenarios_weak_supp}
\end{figure*}

\begin{figure*}[h]
	\centering
	\includegraphics[width=0.9\columnwidth]{{figures/peak_ratio_0.3}.pdf}
	\includegraphics[width=0.9\columnwidth]{{figures/peak_ratio_0.7}.pdf}
	\caption{{\bf Ratio of first and second peak in temperate zone pandemic scenarios, with varying seasonal forcing.} This figure is the same as plotted on the right of Figure \ref{fig:nCov_predictions}, but here seasonal forcing ($\forcing$) is shown at $\forcing=0.3$ (left) and $\forcing=0.7$ (right). Figure \ref{fig:nCov_predictions} uses $\forcing=0.5$.}
	\label{fig:peak_ratio_supp}
\end{figure*}

\begin{figure*}[h]
	\centering
	{\large$R_0=1.4$}
	\includegraphics[width=\textwidth]{{figures/global_3_panel_1.4}.pdf}	{\large$R_0=2.7$}
	\includegraphics[width=\textwidth]{{figures/global_3_panel_2.7}.pdf}
	\caption{{\bf Extended circulation through overlapping epidemics  in variable subpopulations, with varying $R_0$ values.} The same parameter values as in Figure \ref{fig:global_3_panel} are used here, except the average $\langle R_0\rangle$ is 1.4 (top) and 2.7 (bottom). The lower $R_0$ shifts the peak in case numbers to late 2020 and early 2021, while a higher $R_0$ leads to a peak in early-mid 2020.}
	\label{fig:global_3_panel_supp}
\end{figure*}

\newpage

\begin{figure*}[h]
	\centering
	\includegraphics[width=0.8\columnwidth]{{figures/global_endemic_1.5}.pdf}
	\includegraphics[width=0.8\columnwidth]{{figures/global_endemic_3.0}.pdf}
	\caption{{\bf Transition to an endemic seasonal virus, with varying $R_0$ values.} These figures are plotted in the same manner as Figure \ref{fig:endemic}, but with $R_0$ values of 1.5 (left) and 2.7 (right). A lower $R_0$ allows for a relatively smoother transition to seasonality, particularly for Northern temperate regions, while a higher $R_0$ leads to a longer period of comparatively depressed circulation before the establishment of stable seasonality.}
\end{figure*}

\end{document}